A new tidy data structure to support exploration and modeling of temporal data

### 2019-08-22 reviews/response.Rmd

---
title: Response to reviewers
author: JCGS-19-044
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

We thank the reviewers for their helpful feedback which has led to many improvements in the paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.

This manuscript has undergone an overhaul to address reviewer's comments:

1. Section 2.2 (previously 2.3) reviews existing time series and longitudinal data standards in popular statistical software, such as R, Stata, and SAS.
2. We have rewritten Section 3 "Contextual semantics" to give a definition of tsibble in relation to tidy data, and we expand on the four rules of a tsibble in details.
3. We have rewritten Section 4 "Temporal data pipelines" to describe a big picture about each module with the focus of temporal data instead of time series before. We have added Section 4.4 to summarise what a fluent pipeline looks like for a tsibble.
4. We have removed the mosaic plot along with the paragraph from Section 6.1, and added the forecasting example to the smart meter data in Section 6.2 to show tsibble's capability.

Besides these major changes, we have done lots of minor tweaks to improve the manuscript.

@@

> More mini examples embeded throughout the manuscript. Mainly it didn't showcase to me tsibble's capability well enough but understandably this may make the manuscript long and may be better left to vignettes.

TODO.

> Finding a duplicate is a nice touch but I don't understand the significance of the Figure 5 mosaic plot in demonstration for the capability of tsibble. I understand that finding duplicate/wrong entry is made easier by tsibble but besides filtering those duplicate/wrong entries, the data wrangling for mosaic plot is easily done without tsibble. More specifically what is the significance `tsibble::index_by` below? Could the authors explain what is the benefit of explicit defining index at this stage as opposed to explicit specification at a later stage?

We have removed the mosaic plot and the paragraph. The `tsibble::index_by()` helps to update the index and interval when used with `summarise()`. It's a useful adverb to have for intuitive temporal manipulation. But you're right it can happen at a later stage.

@@

> In relation to the above point, perhaps it would be helpful to have a definition of tidy temporal data much like the three rules in tidy data (every row is an observation etc..). E.g. in tsibble the semantics index is a variable with some inherent ordering from past to present.

Thanks for this great suggestion. We've given a definition of "tsibble" following the "tidy data" principles in the beginning of Section 3.

> I hold a similar sentiment regarding the calendar plot in Figure 12. It's a great plot but I don't necessary think this is a plot that show cased the benefit of tsibble.

Yes, the calendar plot just requires a data frame. We've added the forecasting example to back up this plot: how the weather information improves the demand forecasts.

@@

> Figure 11 with missing data was a great demonstration for tsibble and the data times series pipeline shown in Figure 3. It may be helpful to reduce the opacity of the points as there is a lot of overplotting. The top line which lumps a number of customers together is not particularly informative here. Perhaps it is better replaced with a histogram.

TODO.

> Authors may like to add/change `spread`/`gather` verbs to `pivot_longer`/`pivot_wider` with the recent move to deprecate the former in tidyr. Likewise authors may like to consider to replace “long data” to “longer data” and “wide data” to “wider data” as long/wide implies absolute terms. <https://tidyr.tidyverse.org/dev/articles/pivot.html>

We'd like to upgrade `spread`/`gather` verbs to `pivot_longer`/`pivot_wider`. But they are not S3 generics yet, and embedding tsibble into the new vctrs framework will take a long time, which delays the implementation. Conceptually the newly melt columns will be added to the key specification.

@@

> Manuscript does implicitly feel more time series data oriented rather than a general temporal data. E.g. references of time points as index seems inherent from other time series data infrastructure (i.e. zoo and xts) and the presented case studies is not representative of any longitudinal study. This does not mean it should be changed but index to me was not as intuitive to be the temporal variable.

TODO.

> Figure 4 animation doesn't play so authors may like to ensure that the final publication works. Github version works.

The JCGS peer-reviewed version is processed with watermark, which prevents the animation playing.

@@

> Have the authors thought about how to deal with mixed time resolutions for index? E.g. some subjects have exact day recorded; some missing day but month & year recorded; some only date range recorded.

We have added the sentence to Section 3.3, "Although tsibble reports one single interval despite data of fixed or mixed intervals, data of mixed intervals should be organized in separate tsibbles instead for well-tailored analysis and models.", to address the mixed time resolutions.

@@

> The emphasis of the manuscripts is in R but do other languages also tend to define time index implicitly?

We have rewritten Section 2.2 (previously 2.3) to review a broader range of existing data structures in R, Stata, and SAS.

@@

> P7L50 "This diversity and time zone..." suggest rephrase "The tsibble's index column leverages the date-time data type in R to accommodate diverse temporal representations and associated time zone information."

We removed that sentence.

> P8L16 "The identifying variables from which the "key" is constituted remain the same as in the original table with no further tweaks." It's not totally clear to me what is meant here. Is it saying that like with grouping columns (i.e. columns supplied to group_by()), key columns cannot be altered in subsequent analyses? It might be useful to make the comparison to group_by() explicit as it also allows multiple fields. The distinction between group_by() and key() is clear in that the collection of columns that constitute a key must uniquely identify observations (obviously this isn’t the case for grouping variables).

We have re-written Section 3.2 about "key" to be clearer on defining observational units. `group_by()` is a transient operation and not limited to certain variables, but "key" persistently identifies observational units. But "key" can be altered when the observational units are changed during the course of transformation.

@@

> P10L19 "Duplicates signal a data quality..."

Done.

> P12-13 Consider referencing "fable: and possibly "mable" too.

We added the reference to the **fable** package, and removed the mention of mable in Section 4.3.

@@

> P14-15 I had to read section 5.2 a few times before I understood it... I think what threw me was the last paragraph starts with a reference to purrr, but the functions described in the remainder of the paragraph describes functions in the tsibble package. Perhaps move the purrr sentence up to the previous paragraph when functional programming is first mentioned.

Done. We moved **purrr** to the opening sentence of the previous paragraph.

@@

> P15L53 "On the other hand, these tsibble verbs can incorporate many third-party vector functions to step out of the current tsibble zone" perhaps "The extensible design of the tsibble package means that the table verbs can incorporate many third-party vector functions."

Done.

> Fig 5 y-axis doesn't need the overall label "Delayed" or if it stays, should be renamed “Status”. Similarly for the legend label.

We removed the mosaic plot (previously Figure 5).

> Fig 6 Consider replacing gather() with pivot_longer() which is intended to be easier to use particularly for more complex "gathering" operations. <https://tidyr.tidyverse.org/dev/articles/pivot.html>

We'd like to upgrade `spread`/`gather` verbs to `pivot_longer`/`pivot_wider`. But they are not S3 generics yet, and embedding tsibble into the new vctrs framework will take a long time, which delays the implementation. Conceptually the newly melt columns will be added to the key specification.

@@

> Fig 12 remove legend label "hot"

Done.

> P25L19 It's not "A missing piece of the tsibble package" because later in the paragraph it says that custom calendars are easily embedded in the tsibble framework. I suggest rephrasing the first sentence in a more positive light... "The tsibble package can easily be extended to accommodate user-defined calendars." Suggest also referencing at least one package that provides a user-defined calendar (perhaps a market-specific business calendar).

We have re-written the custom calendars part, the last paragraph of Section 7. Conceptually it's easy to embed custom calendars into the tsibble framework, but a general flexible calendar software system is lacking, which delays the implementation. We added the **bizdays** R package as reference.

### 2019-08-23 tsibble.Rmd
added my response to reviewer's comments
done with smart meter revision

Since observational units are embedded, modelling and forecasting for each unit in a tsibble will be accelerated. The tsibble key plays a role of the central transit hub in connecting multiple tables managed by the data, models, and forecasts. This neatly decouples the expensive data copy from downstream summaries, significantly reducing the storage space.

@@

Modeling is crucial to explanatory and predictive analytics, where time series and longitudinal data analysis start to diverge. Nevertheless, tsibble as a model-oriented object can flow into both types of models, with new semantics (index and key) to be internally utilized.

@@

Sensors have been installed in households across major cities in Australia to collect data for the smart city project. One of the trials is monitoring households' electricity usage through installed smart meters in the area of Newcastle over 2010--2014 [@smart-meter]. Data from 2013 have been sliced to examine temporal patterns of customer's energy consumption with **tsibble** for this case study. Half-hourly general supply in kwH have been recorded for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total. Daily high and low temperatures in Newcastle in 2013 provides explanatory variables other than time in a different data table [@newcastle-temp], obtained using the R package **bomrang** [@R-bomrang]. Two data tables might be joined to explore how local weather can contribute to the variations of daily electricity use when needed.

@@

Aggregation across all individuals helps to sketch a big picture of the behavioral change over time, organized into a calendar display (Figure \@ref(fig:calendar-plot)). Each glyph represents the daily pattern of average residential electricity usage every thirty minutes. Higher consumption is indicated by higher values, and typically occurs in daylight hours. Color indicates hot days. The daily snapshots vary depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees Celsius. However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening. This plot illustrates how the tsibble data can easily integrate with other tools and graphics.

@@

@@

@@

A nice feature that the tsibble structure should take into account is to allow user-defined calendars and to respect structurally missing observations. For example, a call center may operate only between 9:00 am and 5:00 pm on week days, and stock trading resumes on Monday straight after Friday. No data available outside trading hours would be labeled as structural missingness, which tsibble currently disregards. Customer calendars can be embedded into the tsibble framework in theory. A few R packages provide functionality to create and manage many specific calendars, such as the **bizdays** package [@R-bizdays] for business days calendars. However, the lack of a generic calendar system delays the implementation. This is left for future work.

>>>>

Since observational units are embedded, modelling and forecasting across units and time in a tsibble will be simplified. The tsibble key plays a role of the central transit hub in connecting multiple tables managed by the data, models, and forecasts. This neatly decouples the expensive data copy from downstream summaries, significantly reducing the storage space.

@@

Modeling is crucial to explanatory and predictive analytics, where time series and longitudinal data analysis start to diverge. Nevertheless, tsibble as a model-oriented object can flow into both types of models, with new semantics (index and key) to be internally utilized, to accelerate modelling.

@@

Sensors have been installed in households across major cities in Australia to collect data for the smart city project. One of the trials is monitoring households' electricity usage through installed smart meters in the area of Newcastle over 2010--2014 [@smart-meter]. Data from 2013 have been sliced to examine temporal patterns of customer's energy consumption with **tsibble** for this case study. Half-hourly general supply in kwH have been recorded for `r format(tsibble::n_keys(elec_ts), big.mark = ",")` customers in the data set, resulting in `r format(NROW(elec_ts), big.mark = ",")` observations in total. Daily high and low temperatures in Newcastle in 2013 provides explanatory variables other than time in a different data table [@newcastle-temp], obtained using the R package **bomrang** [@R-bomrang]. Aggregating the half-hourly energy data to the same daily time interval as the temperature data allows to join the two data tables to explore how local weather can contribute to the variations of daily electricity use and the accuracy of demand forecasting.

@@

Aggregation across all individuals helps to sketch a big picture of the behavioral change over time in the region, organized into a calendar display (Figure \@ref(fig:calendar-plot)) using the **sugrrants** package [@wang-calendarviz-2018]. Each glyph represents the daily pattern of average residential electricity usage every thirty minutes. Higher consumption is indicated by higher values, and typically occurs in daylight hours. Color indicates hot days. The daily snapshots vary depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes predominant driven by the use of air conditioning, especially on hot days with daily average temperature greater than 25 degrees Celsius. However, the winter time (July and August) sees two peaks in a day, which is probably due to heating in the morning and evening.

@@

Lastly, a common practice in the energy sector is load forecasting. Forecasting the demand for the last day of 2013 is undertaken, holding the original energy data as the validation set. The most recent December data is fitted by ARIMA models with and without temporature information using automatic order selection [cite]. The logarithmic transformation is applied to the average demand to ensure the positive forecasts. Figure \@ref(fig:elec-forecast) plots the one-day forecasts obtained from both models against the actual demand within the last week window. ARIMA regressed by average temporatures gives a better fit than the one without, although both tend to underestimate the night demand. The forecasting performance is reported in Table \@ref(tab:elec-accuracy), consistent with the findings in Figure \@ref(fig:elec-forecast).

```{r elec-model}
```

```{r elec-forecast, fig.height = 3, fig.cap = "(ref:elec-forecast-cap)"}
```

(ref:elec-forecast-cap) One-day forecasts generated by ARIMA models with and without temporatures plotted against the actual demand. Both nicely capture the temporal dynamics, but ARIMA with temporatures performs better than the one without.

<!--

@@

-->

```{r elec-accuracy}
```

(ref:elec-accuracy) Accuracy measures to evaluate the forecasting performance between ARIMA models with and without temporatures, using the validation set.

This case study demonstrates the significance of tsibble in lubricating the plumbing of handling time gaps, visualizing, and forecasting in general.

@@

A nice feature that the tsibble structure should take into account is to allow user-defined calendars and to respect structurally missing observations. For example, a call center may operate only between 9:00 am and 5:00 pm on week days, and stock trading resumes on Monday straight after Friday. No data available outside trading hours would be labeled as structural missingness, which tsibble currently disregards. Customer calendars can be embedded into the tsibble framework in theory. A few R packages provide functionality to create and manage many specific calendars, such as the **bizdays** package [@R-bizdays] for business days calendars. However, a generic flexible calendar system is lacking, that delays the implementation. This is left for future work.
